Fail Faster

### 2025-03-19 paper/paper.tex
nit
Making some progress on section 2.1

\settopmatter{printfolios=false,printccs=false,printacmref=false}
\setcopyright{none}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.17}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{xspace}
\lstset{
  mathescape=true,
  frame=none,
  xleftmargin=10pt,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  % language=haskell,
  keepspaces=true,
  tabsize=2,
  emphstyle={\bf},
  % commentstyle=\it\color{dkgreen},
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\bfseries\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  % morecomment=[l]\%,
  % moredelim=**[is][\color{dkgreen}]{@}{@}
}

@@

\subsection{Background: Monadic Generator DSLs}
\jwc{Usual spiel about pbt and generators: here's a monadic generic generator
library (but we're using BQ to illustrage), it has return and bind, this is what
they're for.}
\jwc{You often want generators to generate for sparse precoditions
so you can exercise a property reasonably well}

Property Based Testing was popularized by the QuickCheck library \cn{} in
Haskell.  A core element of PBT as introduced in that paper is \emph{monadic
generators}, an embedded DSL for writing the random data generators that drive
the PBT process.  This DSL gives developers significant control over exactly
what sorts of values are generated and what their distribution is. Both of these
axes of control are important to adequately exercise the system under test and
find bugs.  Indeed, the PBT literature \cn{} has repatedly demonstrated that the
particular generator used to test a property has an enormous impact on how well
PBT actually works \jwc{less derogatory phrasing}, and much research has gone
into finding clever ways to generate data \cn{}.

\begin{figure}
\begin{lstlisting}
module Bq : sig
  type 'a t
  val return : 'a -> 'a t
  val bind : 'a t -> ('a -> 'b t) -> 'b t
  val gen_int : int -> int -> int t

@@

  val fixed_point : ('a t -> 'a t) -> 'a t
  $\dots$
end
\end{lstlisting}
\caption{API of the \bq{} Generator DSL}
\label{fig:bq-api}
\end{figure}

A standard generator DSL consists of a monadic interface, plus combinators for sampling random numbers, making (weighted) random choices,
and controlling the size of generated values. Figure~\ref{fig:bq-api} shows a selection of functions from the API of th \bq{} generator
DSL.

Because we have to choose \emph{some} syntax for this paper, we use \bq{}'s as our primary running generator DSL. However,
this is a standin for a generic generator DSL, since they basicalyl all look like this one.

The ``monadic interface'' of a generator DSL consists of \texttt{return} and \texttt{bind}.
The generator \texttt{return x} is the constant generator, always generating the value \texttt{x}.
Running a generator \texttt{bind g k} runs the generator \texttt{g}, producing a value \texttt{a},
and then runs the generator \texttt{k a}. Monadic \texttt{bind} allows for dependencies.

The \texttt{gen\_int a b} generator samples an integer uniformly in the inclusive range from \texttt{a} to \texttt{b}.

\jwc{\texttt{gen\_int} here is really \texttt{int\_uniform\_inclusive}. I also think we should completely omit named arguments in the paper so we don't confuse with splices.}

Figure ... shows an example of a generator written with these combinators. It first samples an int between $0$ and $100$, 
names it $x$, samples another between $0$ an $x$, and then returns the pair of them both.
Most languages in which monadic APIs are common expose some sort of sytnactic sugar for them --- this is true of OCaml and Scala. 
OCaml's looks like \texttt{let\%bind x = e in e'}, which desguars to \texttt{...}. Figure~\jwc{...} also shows this generator written with OCaml's syntacitc sugar\footnote{It's actually Jane's}.

\begin{figure}

\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  Bq.bind (Bq.gen_int 0 100) (fun x ->
    Bq.bind (Bq.gen_int 0 x) (fun y ->
      Bq.return (x,y)
  ))
\end{lstlisting}

\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  let%bind x = Bq.gen_int 0 100 in
  let%bind y = Bq.gen_int 0 x in
  return (x,y)
\end{lstlisting}
\end{figure}

@@

Strategies for BST and STLC are unchanged from \secref{subsection:benchmarking}.
To ensure a fair comparison, we initialize each of these generators with the same random seed, ensuring that they produce identical sequences of values. We evaluate each strategy on all tasks using a 60-second timeout. 
If a bug is not found within this limit, the task is considered unsolved. 
Tasks where all strategies fail to find the bug are excluded from our dataset, 
as they provide no basis for comparison. 
Similarly, we exclude tasks where the BQ strategy completes in under 5ms,
 as such negligible runtimes do not yield meaningful insights into relative performance. These filters ensure that reported speedups are based on cases where optimization has a measurable impact.

>>>>

\settopmatter{printfolios=false,printccs=false,printacmref=false}
\setcopyright{none}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.17}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{wrapfig}
\lstset{
  mathescape=true,
  frame=none,
  xleftmargin=10pt,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  % language=haskell,
  keepspaces=true,
  tabsize=2,
  emphstyle={\bf},
  % commentstyle=\it\color{dkgreen},
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\bfseries\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  % morecomment=[l]\%,
  % moredelim=**[is][\color{dkgreen}]{@}{@}
}

@@

\subsection{Background: Monadic Generator DSLs}

Since property-based testing was first introduced in Haskell's QuickCheck
library~\cite{claessenQuickCheckLightweightTool2000}, {\em generators} have been
a key element of the PBT process. While properties are critical for correctly
expressing a developer's intent for a function, the success of the testing
ultimately rests on how well the random inputs to that property exercise the
system under test. QuickCheck's solution to this problem, which has been copied
dozens of frameworks across as many languages, is to provide a domain-specific
language (DSL) for writing random data generators that gives developers complete
control over their distribution of test inputs.

\begin{wrapfigure}{r}{.5\textwidth}
\begin{lstlisting}
module Bq : sig
  type 'a t
  val gen_int : int -> int -> int t
  val return : 'a -> 'a t
  val bind : 'a t -> ('a -> 'b t) -> 'b t

@@

  val fixed_point : ('a t -> 'a t) -> 'a t
  $\dots$
end
\end{lstlisting}
\caption{Some functions from the API of the \bq{} generator DSL.}
\label{fig:bq-api}
\end{wrapfigure}

While many DSL designs are possible, the standard design for QuickCheck-style
generators is via an embedded {\em monadic} language. For the purposes of this
paper, we use the sytnax from \bq, which is presented in
Figure~\ref{fig:bq-api}. The library provides a series of basic generators, for
example ones like \texttt{gen\_int} for random integers in a range, along with
the ``monadic interface'' consisting of \texttt{return} and \texttt{bind}.  The
generator \texttt{return x} is the constant generator, always generating the
value \texttt{x}. Running a generator \texttt{bind g k} runs the generator
\texttt{g}, producing a value \texttt{a}, and then runs the generator \texttt{k
a}.
\iflater
\jwc{\texttt{gen\_int} here is really \texttt{int\_uniform\_inclusive}. I also
think we should completely omit named arguments in the paper so we don't confuse
with splices.}
\fi

Together, these three functions are the bare minimum for constructing arbitrary
random data generators: \texttt{gen\_int} provides a base source of randomness,
and \texttt{return} and \texttt{bind} allow generators to be composed to create
larger, more complex generators. Figure~\ref{fig:simple-bind} shows a generator
built with these operations; it first samples an int between $0$ and $100$,
names it $x$, samples another between $0$ an $x$, and then returns the pair of
them both. In practice, PBT generator libraries also include other functions
that make generator construction easier, including functions like
\texttt{weighted\_union} for making a weighted choice between different
generators as well as functions for managing the sizes of generated values and
combinators for recursion.

\begin{figure}[h]
  \begin{subfigure}{.49\textwidth}
\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  Bq.bind (Bq.gen_int 0 100) (fun x ->
    Bq.bind (Bq.gen_int 0 x) (fun y ->
      Bq.return (x,y)))
\end{lstlisting}
\caption{A simple generator using \texttt{bind} explicitly.}\label{fig:simple-bind}
  \end{subfigure}
  \begin{subfigure}{.49\textwidth}
\begin{lstlisting}
let int_pair : (int * int) Bq.t =
  let%bind x = Bq.gen_int 0 100 in
  let%bind y = Bq.gen_int 0 x in
  return (x,y)
\end{lstlisting}
\caption{An equivalent generator using macros for \texttt{bind}.}\label{fig:simple-macro}
  \end{subfigure}
  \caption{Simple monadic generators for a pair of ordered integers.}\label{fig:simple-pair}
\end{figure}

Most languages in which monadic APIs are common expose some sort of syntactic
sugar for them. In OCaml\footnote{OCaml actually has a few versions of syntax
for monadic APIs; this is the one provided by Jane Street's libraries.}, this
looks like \texttt{let\%bind x = e in e'}, which is a macro that desugars to
\texttt{bind e (fun x => e')}. Figure~\ref{fig:simple-macro} shows the same
generator as before, now written with macros.

@@

Strategies for BST and STLC are unchanged from \secref{subsection:benchmarking}.
To ensure a fair comparison, we initialize each strategy with the same random seed, ensuring that they produce identical sequences of values. We evaluate each strategy on all tasks using a 60-second timeout. 
If a bug is not found within this limit, the task is considered unsolved. 
Tasks where all strategies fail to find the bug are excluded from our dataset, 
as they provide no basis for comparison. 
Similarly, we exclude tasks where the BQ strategy completes in under 5ms,
 as such negligible runtimes do not yield meaningful insights into relative performance. These filters ensure that reported speedups are based on cases where optimization has a measurable impact.
